---
title: "1-Make an Histogram"
subtitle: "“R, in itself, is an attempt to bring order out of chaos.”"
author: "Otho Mantegazza<br>`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: css/remark.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---


```{r, include=FALSE}
knitr::opts_chunk$set(dev = "svg")
```

# The Rise of Private Space Agencies 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 5, fig.width=10}
launches <- 
  read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-15/launches.csv")

launches %>% 
  ggplot(aes(x = launch_year,
             fill = agency_type)) +
  geom_bar(alpha = .8, width = 1.02) +
  labs(x = "",
       y = "Space Launches [n]",
       fill = "Agency Type:") +
  scale_fill_manual(values = c("#27A6D3", "#FF6C0D", "#100089"),
                    guide = guide_legend(keyheight = .7, keywidth = 2)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")
```

---

class: blueblue, middle

.verybig[So what?]

--

.big[Let's start from the beginning...]


---

class: blueblue, middle, right

.big[Let's **read** the data...]

--

.big[...into an R **object**...]

--

.big[...and store it into a **variable**.]

---

# R understands you

Write something in the R console:

```{r}
2 # numbers
```

--

```{r}
"string of text must be quoted" # text
```

--

```{r}
2 + 2 # any mathematical operation

3^2
```

---

# R understands you

you can group together numbers

```{r}
c(1, 2, 5, 6, 8, 1, 45, .2, -1) # or numbers
```

--

operations are vectorized.

```{r}
c(1, 2, 5, 6, 8, 1, 45, .2, -1)*2 # also for numbers
```

---

# R understands you

The same for letters.

```{r}
c("you", "can", "group", "together", "strings", "of", "text")
```

--

```{r}
paste(c("you", "can", "group", "together",
        "strings", "of", "text"),
      collapse = " ") # operation are vectorized
```


---

class: blueblue, middle

.verybig[When you type something in R, R guesses what it is.]

---

# Everything in R is an object

And objects have types

--

```{r}
class(1)
```

--

```{r}
class("ciao")
```

--

```{r}
class(FALSE)
```

--

```{r}
class(mean)
```

--

exercise: put together objects of different classes with `c(..., ...)` and see what happens.

```{r}
class(c(2,3,5,7,4)) # collection of numbers or strings
```

--

```{r}
mean # functions
```

---

class: blueblue, middle

.big[You can discover the type of an object with `class()`]

---

Also data are objects...

```{r}
mpg
```

--

```{r}
class(mpg)
```

