---
title: "2 The Tidyverse"
subtitle: "“R does not reproduce what we see. It makes us see.”"
author: "Otho Mantegazza<br>`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: css/remark.css
    # chakra: libs/remark-js-latest-min.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
source(file = here::here("slides/libs/xaringan_reveal_parentheses_balanced.R"))
```


# The Tidyverse

.center[

</p>
<a href="https://www.tidyverse.org/" class="imagelink">
<img src="img/SVG/tidyverse.svg" alt="hex-tidyverse" height="400" width="400"></a>
<p>

]

---

class: blueblue, middle

.verybig[The Tidyverse is a collection of packages for Data Science]

---


class: blueblue, middle

.big[.right[Let's practice it!]]

---

# NYC Squirrel Census

.middle[

.pull-left[

Because:

- It is large enough to provide a challenge.
- It is tidy and detailed,
- It stores quantitative and categorical variables,
- it stores spatial variables,
- it stores time variables.

]

.pull-right[


.center[

</p>
<a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-10-29" class="imagelink">  data here
<img src="img/squirrel-svgrepo-com.svg" alt="hex-tidyverse" height="400" width="400"></a>
<p>

]
]
]

---

# NYC Squirrel Census

`r emo::ji("squirrel")`

```{r, message=FALSE}
library(tidyverse)
```


```{r, message=FALSE}
squirrels_url <- paste0("https://raw.githubusercontent.com/rfordatascience/tidytuesday/",
                        "master/data/2019/2019-10-29/nyc_squirrels.csv")

squirrels <- read_csv(squirrels_url)
```


---


# NYC Squirrel Census `r emo::ji("tree")`

```{r}
squirrels
```

---

# NYC Squirrel Census

```{r squirrel-pipe-1, echo=FALSE, eval=FALSE}
squirrels %>% 
  select(lat_long, primary_fur_color, location) %>% 
  filter(primary_fur_color == "Cinnamon") %>% 
  drop_na(location)
  
```

`r apply_reveal("squirrel-pipe-1")`
